---
title: "P7_Initial_Analysis"
author: "Kevin Rose"
date: "Last updated on: 7/6/2022"
output: 
  html_document: 
    toc: yes
    toc_float: yes
    theme: readable
    fig_width: 10
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is the initial analysis of the OMC-rich P7 scRNA-seq dataset using the Seurat pipeline. Steps included in the initial analysis:  

1) Loading in the P7 dataset:
  You will need the following cellranger outputs (filtered) which includes:
    1) barcodes.tsv  
    2) features.tsv  
    3) matrix.mtx  
2) Quality Control  
3) Normalization, identification of highly variable genes, and scaling the data
4) Linear Dimensionality Reduction via PCA
5) Clustering (graph-based)
6) Visualization via UMAP

## Loading Data into Seurat

```{r, Data_Upload, warning = FALSE, message = FALSE}
#Libraries needed for analysis
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)

#P7 Cochlea from CellRanger
P7.data <- Read10X(data.dir = "./CellRanger/")
# Initialize the Seurat object with the raw (non-normalized data).
P7 <- CreateSeuratObject(counts = P7.data, project = "P7", min.cells = 3, min.features = 200)
rm(P7.data)
```


## Quality Control Steps

```{r, QC}
#Quality Control
#Mitochondrial Percentage
P7[['percent.mt']] <- PercentageFeatureSet(P7, pattern = '^mt-')
#Visualize QC metrics - VlnPlots
VlnPlot(P7, cols = 'cyan4', features = c('nFeature_RNA', 'nCount_RNA', 'percent.mt'), ncol = 3)
#QC ScatterPlots
plot1 <- FeatureScatter(P7, cols = 'cyan4', feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(P7, cols = 'cyan4', feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2
#cutoffs = percent_mt = 30% | nFeatures (genes) = 200 - 7500
P7 <- subset(P7, subset = nFeature_RNA > 200 & nFeature_RNA < 7500 & percent.mt < 30)
```

## Normalization, identification of highly variable genes, scale data

```{r, Norm_variable_scale, warning = FALSE, message = FALSE}
#Normalization by total expression with 10,000 scale factor
P7 <- NormalizeData(P7, normalization.method = "LogNormalize", scale.factor = 10000)

#Identify top 2000 Highly Variable Genes
P7 <- FindVariableFeatures(P7, selection.method = "vst", nfeatures = 2000)
# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(P7), 10)
# plot variable features with and without labels
plot1 <- VariableFeaturePlot(P7, cols = c('gray75', 'cyan4'))
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot2

#Scale Data
all.genes <- rownames(P7)
P7 <- ScaleData(P7, features = all.genes)
```

## Linear Dimensional Reduction (PCA)

```{r, PCA, warning = FALSE, message = FALSE}
#Linear Dimensional Reduction
P7 <- RunPCA(P7, features = VariableFeatures(object = P7))
#Visualize the first 2 PCs
VizDimLoadings(P7, dims = 1:2, reduction = "pca")
DimPlot(P7, reduction = "pca", cols = 'cyan4')
```

## Clustering

```{r, Cluster, warning = FALSE, message = FALSE}
#Determine the dimensionality of the dataset
P7 <- JackStraw(P7, num.replicate = 40)
P7 <- ScoreJackStraw(P7, dims = 1:20)
JackStrawPlot(P7)
ElbowPlot(P7)

#Cluster cells with 0.5 resolution and 15 PCs
P7 <- FindNeighbors(P7, dims = 1:15)
P7 <- FindClusters(P7, resolution = 0.5)
```

## Visualization via UMAP

```{r, UMAP, warning = FALSE, message = FALSE}
#Run non-linera dimensional reduction (UMAP)
P7 <- RunUMAP(P7, dims = 1:20)

P7 = readRDS("./RDS_Files/Pou3f4_WT_P7.rds") #Loading in originally computed rds file so UMAP matches manuscript. UMAP generation has some randomness to it.
DimPlot(P7, reduction = 'umap', pt.size = 2, label = TRUE)
```

